Benchmarking neuromorphic systems with Nengo

### 2015-07-29 paper/paper.tex
Move supplementary material to main paper
There's plenty of room for it, so why not!

\usepackage{url,lineno}
\linenumbers
\usepackage{inconsolata}

@@

An example script showing how Nengo can
express a functional network
with 500~neurons and 6~connections
(representing 50,000~connection weights
and 100~direct current injection sites)
in under 20~lines of code
can be seen in Supplementary Figure~2.

@@

Nengo's test suite employs the \texttt{pytest}
testing framework.\footnote{Available at \url{http://pytest.org/}.}
\texttt{pytest} enables
expressive testing of pluggable components
(such as backends)
through what are called ``test fixtures.''
Fixtures are exposed to the test suite
as arguments that can be provided
to any test function.
\texttt{pytest} inspects the function signature,
and passes an appropriate value
to the function.
This allows for boilerplate code
to be specified once,
and run for any test that might use it.
A demonstration of how these fixtures
are used in test scripts
can be seen in Supplementary Figure~2.

@@

\paragraph{\texttt{analytics\_data}}
enables comparative testing between
two or more runs of the same test.
The runs might represent multiple versions
of the same backend---one before a
speedup and one after---or multiple
entirely different backends.
The \texttt{analytics\_data} fixture
provides the results from those two
test runs to a single test function,
allowing for comparisons
between the two results saved
with the \texttt{analytics} fixture.
See Supplementary Figure~12Appendix~\ref{App:01} for a
concrete example of using the
\texttt{analytics\_data} fixture.

@@

Each test has several parameters
that can be varied
in order to benchmark a wide variety of cases
(see Supplementary Figure~2 for an example).
We test each model with parameters
typically used in large-scale models,
but make these parameterized models
available at \url{https://github.com/ctn-archive/bekolay-fnme2015}
for those who wish to explore additional cases.

@@

The versions of software
used for benchmarking can be found in
Supplementary Table~1.

@@

Interestingly, while removing passthrough nodes
results in significantly slower run times
for the reference backend,
it speeds up the OpenCL and SpiNNaker backends slightly.
This result indicates that passthrough nodes
add to the overhead incurred by
SpiNNaker and OpenCL.
Other backends also incur overhead,
especially when data is recorded
through probes.
We investigated the possibility
that the overhead involved in probes
dominated the run time for some backends.
However, we found that
removing probes from these models
made very little impact on
the total time taken in a \texttt{run} call
(see Supplementary Figure~1).

@@

All of the software packages discussed in this paper
are available online,
except for the Brainstorm backend,
which has not yet been publicly released
(see Supplementary Table~1).
All of the benchmarks presented here,
as well as the scripts used to generate
the presented figures,
are available at
\url{https://github.com/ctn-archive/bekolay-fnme2015}.
The results of running the benchmarks
are not included in the repository;
however, they can be downloaded separately at
\url{http://dx.doi.org/10.6084/m9.figshare.1496569}
to replicate the figures in this paper.

@@

@@

@@

>>>>

\usepackage{url,lineno}
\linenumbers
\usepackage{inconsolata}
\usepackage{listings}
\lstset{language=Python}
\lstset{basicstyle=\ttfamily,breaklines=true}

@@

An example script showing how Nengo can
express a functional network
with 500~neurons and 6~connections
(representing 50,000~connection weights
and 100~direct current injection sites)
in under 20~lines of code
can be seen in Appendix~A.

@@

Nengo's test suite employs the \texttt{pytest}
testing framework.\footnote{Available at \url{http://pytest.org/}.}
\texttt{pytest} enables
expressive testing of pluggable components
(such as backends)
through what are called ``test fixtures.''
Fixtures are exposed to the test suite
as arguments that can be provided
to any test function.
\texttt{pytest} inspects the function signature,
and passes an appropriate value
to the function.
This allows for boilerplate code
to be specified once,
and run for any test that might use it.
A demonstration of how these fixtures
are used in test scripts
can be seen in Appendix~A.

@@

\paragraph{\texttt{analytics\_data}}
enables comparative testing between
two or more runs of the same test.
The runs might represent multiple versions
of the same backend---one before a
speedup and one after---or multiple
entirely different backends.
The \texttt{analytics\_data} fixture
provides the results from those two
test runs to a single test function,
allowing for comparisons
between the two results saved
with the \texttt{analytics} fixture.
See Appendix~A for a
concrete example of using the
\texttt{analytics\_data} fixture.

@@

Each test has several parameters
that can be varied
in order to benchmark a wide variety of cases
(see Appendix~A for an example).
We test each model with parameters
typically used in large-scale models,
but make these parameterized models
available at \url{https://github.com/ctn-archive/bekolay-fnme2015}
for those who wish to explore additional cases.

@@

The versions of software
used for benchmarking can be found in
Table~\ref{Tab:02}.

@@

Interestingly, while removing passthrough nodes
results in significantly slower run times
for the reference backend,
it speeds up the OpenCL and SpiNNaker backends slightly.
This result indicates that passthrough nodes
add to the overhead incurred by
SpiNNaker and OpenCL.
Other backends also incur overhead,
especially when data is recorded
through probes.
We investigated the possibility
that the overhead involved in probes
dominated the run time for some backends.
However, we found that
removing probes from these models
made very little impact on
the total time taken in a \texttt{run} call
(see Figure~\ref{fig:07}).

@@

All of the software packages discussed in this paper
are available online,
except for the Brainstorm backend,
which has not yet been publicly released
(see Table~\ref{tab:02}).
All of the benchmarks presented here,
as well as the scripts used to generate
the presented figures,
are available at
\url{https://github.com/ctn-archive/bekolay-fnme2015}.
The results of running the benchmarks
are not included in the repository;
however, they can be downloaded separately at
\url{http://dx.doi.org/10.6084/m9.figshare.1496569}
to replicate the figures in this paper.

@@

\clearpage

\section*{Appendix A: Chained Communication Channel model}

\begin{lstlisting}
import numpy as np
import nengo
from nengo.utils.numpy import rmse

def test_cchannelchain(Simulator, analytics, plt, rng, seed):
    # Parameters that can be varied to investigate extreme cases
    dims = 2
    layers = 5
    n_neurons = 100
    synapse = nengo.Lowpass(0.01)

    with nengo.Network(seed=seed) as model:
        hypersphere = nengo.dists.UniformHypersphere()
        value = hypersphere.sample(dims, 1, rng=rng).ravel()
        stim = nengo.Node(value)

        ens = [nengo.Ensemble(n_neurons, dimensions=dims)
               for _ in range(layers)]

        nengo.Connection(stim, ens[0])
        for i in range(layers - 1):
            nengo.Connection(ens[i], ens[i+1], synapse=synapse)

        p_input = nengo.Probe(stim)
        p_outputs = [nengo.Probe(ens[i], synapse=synapse)
                     for i in range(layers)]

    sim = Simulator(model)
    sim.run(0.5)

    for p_output in p_outputs:
        plt.plot(sim.trange(), sim.data[p_output])
    plt.plot(sim.trange(), sim.data[p_input], color='k', linewidth=1)
    plt.ylabel('Decoded output')
    plt.xlabel('Time (s)')

    last = p_outputs[-1]
    decoding_rmse = rmse(value, sim.data[last][sim.trange() > 0.4]
    analytics.add_data('rmse', decoding_rmse)

def test_compare_cchannelchain(analytics_data, plt):
    rmses = [d['rmse'] for d in analytics_data]
    plt.bar(np.arange(len(rmses)), rmses, align='center')
    plt.ylabel("RMSE")
\end{lstlisting}

@@

\begin{table}[!ht]
\textbf{\refstepcounter{table}\label{Tab:02} Table \arabic{table}.}{
  Software versions used for benchmarking.
  The PyPI name is the unique identifier used for installing the Python package
  through the Python Package Index (PyPI). Packages can be found by
  visiting \url{https://pypi.python.org/pypi/<PyPI name>}.}

\processtable{ }
{\begin{tabular}{lcc}\toprule
 Package & PyPI name & Version number\\\midrule
 Nengo & \texttt{nengo} & Development version, commit \texttt{7d2d24145} \\
 Distilled backend & \texttt{nengo\_distilled} & 0.1.0 \\
 OpenCL backend & \texttt{nengo\_ocl} & 0.1.0 \\
 Brainstorm backend & Not available on PyPI & Development version, commit \texttt{ac3cfa708} \\
 SpiNNaker backend & \texttt{nengo\_spinnaker} & 0.2.4 \\
 NumPy & \texttt{numpy} & 1.8.1 \\\botrule
\end{tabular}}{}
\end{table}

@@

\begin{figure}[!ht]
\begin{center}
  \includegraphics[width=85mm]{fig7}
\end{center}
\textbf{\refstepcounter{figure}\label{fig:07} Figure \arabic{figure}.}
       {Run time for each model with probes removed on each backend.
         Each bar represents the mean run time across 100 instances
         of each model;
         error bars are 95\% confidence intervals.
         Run time is measured relative to real time;
         e.g., a value of two indicates that the model ran twice as slow
         as real time.
         BG Sequence * refers to the basal ganglia sequence model
         in which passthrough nodes have been removed.}
\end{figure}
